Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-01-14 Manuscript/Manuscript.Rmd
Integrate comments last meeting

~~Research into any field with a hierarchical or clustered nature of observations may yield multilevel data. In the typical case, individuals are nested within groups, but there are many different types of multilevel data. In the medical field, clustering occurs at e.g., the hospital or center level in registry data, or at the study-level in meta-analyses (IPDMA). In the social sciences and official statistics we can find clustering e.g. at the country-level, or as imposed by a multi-stage sampling design. For the sake of legibility, we will refer to the grouping variable as 'cluster', and the grouped variable as '(sample) unit' throughout this paper. And, for reasons of brevity, we only discuss clustering between units, not within units (e.g., timeseries or longitudinal data). Also, add that we'll only discuss two levels, not more?~~ Add IPDMA back in: patient level data from multiple studies

In many contemporary data analysis efforts, some form of hierarchical or clustered structure is recorded. Ignoring such structures may be harmful to the inferences and can yield biased results. On the other hand, analyzing such multilevel data requires specialized techniques that take the clustered structure into account. Imagine a case where cross-level interactions between unit-level variables and cluster-level variables are present. The cluster to which a unit belongs may then influence the unit-level observations - and vice versa - for each of the units that make up the cluster [@hox17]. These relations can and should be taken into account when developing analysis models for multilevel data for the simple reason that groups of observations share some common variance.

@@

>>>>

In many contemporary data analysis efforts, some form of hierarchical or clustered structure is recorded.^[Add IPDMA back in: patient level data from multiple studies.] Ignoring such structures may be harmful to the inferences and can yield biased results. On the other hand, analyzing such multilevel data requires specialized techniques that take the clustered structure into account. Imagine a case where cross-level interactions between unit-level variables and cluster-level variables are present. The cluster to which a unit belongs may then influence the unit-level observations - and vice versa - for each of the units that make up the cluster [@hox17]. These relations can and should be taken into account when developing analysis models for multilevel data for the simple reason that groups of observations share some common variance.

@@

<!-- "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix" (Little and Rubin, p. 8). -->
